---
title: "Data Cleaning of Dirty Data"
author: "David Otoosakyi"
date: "October 22, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
This task is aimed at cleaning a messy dataset that was extracted from a database but in excel format.

## Importing the dataset
```{r}
#loading the required library
library(tidyverse) #for data wrangling
library(readxl) #for reading excel data
```

```{r}
#importing the dataset
dirty_data <- read_xlsx("dirty data.xlsx")
head(dirty_data)
tail(dirty_data)
```

```{r}
#chekcing the data dimension of the dataset
dim(dirty_data)
```


```{r}
#checking the structure of the data
str(dirty_data)
```
The above shows that the `transaction` column contains som complex records.

```{r}
#checking the data type
class(dirty_data) 
```

```{r}
#changing structure (class) of the dataset to dataframe
class(dirty_data) = 'data.frame'
```

```{r}
class(dirty_data)
view(dirty_data)
```

```{r}
glimpse(dirty_data)
```

```{r}
#subsetting the `transaction` column
dirty_transaction <- dirty_data[3]
head(dirty_transaction)
```

```{r}
View(dirty_transaction)
```

The following can be extracted from the `transcation` column:
1. **tx_id**
2. **tx_date**
3. **tx_amount**
4. **tx_mode**
5. **tx_channel**
6. **bank_name**
7. **dob**

## Seperating the `transaction` column

```{r}
dirty_transaction <- separate(data = dirty_transaction, col = transaction, into = c("tx_id", "tx_date", "tx_amount","tx_mode","tx_channel","bank_name","dob"), sep = ",") 
view(dirty_transaction)
```

## Eliminating header names appearing in each column entries
```{r}
#loading the required package
library(stringr)
```

```{r}
#tx_id
dirty_transaction$tx_id<-str_sub(dirty_transaction$tx_id, start=12, end = -2)
view(dirty_transaction$tx_id)
```

```{r}
#tx_date
dirty_transaction$tx_date<-str_sub(dirty_transaction$tx_date, start=14, end = -2)
view(dirty_transaction$tx_date)
```

```{r}
#tx_amount
dirty_transaction$tx_amount<-str_sub(dirty_transaction$tx_amount, start = 14)
head(dirty_transaction$tx_amount)
```

```{r}
#trimming off whitespaces
dirty_transaction$tx_amount <- str_trim(dirty_transaction$tx_amount)
head(dirty_transaction$tx_amount)
```

```{r}
#tx_mode
dirty_transaction$tx_mode <- str_sub(dirty_transaction$tx_mode, start = 14, end = -2)
view(dirty_transaction$tx_mode)
```

```{r}
#tx_channel
dirty_transaction$tx_channel <- str_sub(dirty_transaction$tx_channel, start = 17, end = -2)
view(dirty_transaction$tx_channel)
```

```{r}
#bank_name
dirty_transaction$bank_name <- str_sub(dirty_transaction$bank_name, start = 16, end = -2)
view(dirty_transaction$bank_name)
```

```{r}
#dob
dirty_transaction$dob <- str_sub(dirty_transaction$dob, start = 11, end = -4)
view(dirty_transaction$dob)
```

```{r}
#viewing the well separated "dirty_transaction" dataframe
view(dirty_transaction)
```

## Merging the treated datasets
```{r}
dirtydata_separated <- data.frame(dirty_data, dirty_transaction)
view(dirtydata_separated)
```

```{r}
#dropping the "transaction" column
dirtydata_separated <- dirtydata_separated[,-3]
head(dirtydata_separated)
```

```{r}
#checking the structure of the separated data
str(dirtydata_separated)
```

The above result shows that `tx_date()`, `tx_amount()` and `dob()` are having wrong data type. We'll therefore correct this.

### Converting `tx_amount` to `integer`
```{r}
#converting "tx_amount" to integer
dirtydata_separated$tx_amount <- as.integer(dirtydata_separated$tx_amount)
class(dirtydata_separated$tx_amount)
```

### Converting `tx_date` and `dob` to `date-time`
```{r}
#tx_date
dirtydata_separated$tx_date <- as.POSIXct(dirtydata_separated$tx_date, format="%Y-%m-%dT%H:%M")
str(dirtydata_separated)
```

```{r}
glimpse(dirtydata_separated)
```

We now have our`tx_date` as a `date-time` object.

```{r}
#dob
dirtydata_separated$dob <- as.POSIXct(dirtydata_separated$dob,format="%Y-%m-%d %H:%M:%S")
glimpse(dirtydata_separated)
```

Our `dob` column is also in date-time format.

```{r}
view(dirtydata_separated)
```

#Exporting the treated data
```{r}
library(writexl)
```

```{r}
write_excel_csv(dirtydata_separated, "C:/Users/SugarIT/Documents/Data Cleaning of Dirty Data/separated data.csv", col_names = TRUE)
```

#Filling the `NA`s in `dob` with `Sys.Date()`
```{r}
#replicating the file
clean_data <- dirtydata_separated
glimpse(clean_data)
```

```{r}
#filling the missing data
clean_data$dob[is.na(clean_data$dob)] = Sys.Date()
```

```{r}
#verifying the change
glimpse(clean_data)
```

```{r}
write_excel_csv(clean_data, "C:/Users/SugarIT/Documents/Data Cleaning of Dirty Data/clean data.csv", col_names = TRUE)
```

